Checking Invariant Confluence, In Whole or In Parts \thanks{ The original version of this paper is entitled ``Interactive Checks for Coordination Avoidance'' and was published in PVLDB Vol.\ 12. }

### 2020-05-22 doc/sigmod_paper/sections/abstract.tex
Small fixes.

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we
use this condition to design a general-purpose interactive
\invariantconfluence{} decision procedure. We then take a step beyond
\invariantconfluence{} and introduce a generalization of
\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination. We implemented these formalisms in a prototype called Lucy and
found that our decision procedures efficiently handle common real-world
workloads including foreign keys, escrow transactions, and more.
\end{abstract}

>>>>

In this paper, we establish conditions under which a commonly used sufficient
condition for \invariantconfluence{} is both necessary and sufficient, and we
use this condition to design a general-purpose interactive
\invariantconfluence{} decision procedure. We then take a step beyond
\invariantconfluence{} and introduce a generalization of
\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination. We implement these formalisms in a prototype called Lucy and
find that our decision procedures efficiently handle common real-world
workloads including foreign keys, escrow transactions, and more.
\end{abstract}

### 2020-05-22 doc/sigmod_paper/sections/introduction.tex
Cleaned up commented out text.
Small fixes.

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed transactions
(e.g., \cite{bernstein1981concurrency,mohan1986transaction}) or state machine
replication (e.g., \cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search}) makes the application designer's life
very easy. To the developer, a strongly consistent system behaves exactly like
a single-threaded system running on a single node, so reasoning about the
behavior of the system is simple~\cite{herlihy1990linearizability}.
Unfortunately, strong consistency is at odds with performance. The CAP theorem
and PACELC theorem tell us that strongly consistent systems suffer from higher
latency at best and unavailability at worst~\cite{gilbert2002brewer,
brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
models like eventual consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak
consistency models strip an application developer of one of the earliest and
most effective tools that is used to reason about the execution of programs:
application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
if every transaction executing in a weakly consistent system individually
maintains an application invariant, the system as a whole can produce
invariant-violating states.

@@

Unfortunately, to date, the task of identifying whether or not an object
actually is \invariantconfluent{} has remained an exercise in human proof
generation. Bailis et al.\ manually categorized a set of common objects,
transactions, and invariants (e.g.\ foreign key constraints on relations,
linear constraints on integers) as \invariantconfluent{} or not. Hand-written
proofs of this sort are unreasonable to expect from programmers. Ideally we
would have a general-purpose program that can automatically determine
\invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
to make \invariantconfluence{} checkable:} to design a general-purpose
\invariantconfluence{} decision procedure, and implement it in an interactive
system.

@@

In this paper, we take a step back and study the underlying reason \emph{why}
\invariantclosure{} is not necessary for \invariantconfluence{}.
%
% We find that \invariantconfluence{} is fundamentally a reachability property,
% checking whether certain values of our replicated object can actually be
% obtained in the execution of a system. We discover that unreachable, yet
% invariant satisfying, states cause \invariantclosure{} to be an unnecessary
% condition.
%
Using this understanding, we construct a set of modest conditions under which
\invariantclosure{} and \invariantconfluence{} are in fact \emph{equivalent},
allowing us to reduce the problem of determining \invariantconfluence{} to that
of determining \invariantclosure{}.
%
Then, we use these conditions to design a general-purpose interactive
\invariantconfluence{} decision procedure.

% Then, we use these conditions to design a general-purpose interactive
% \invariantconfluence{} decision procedure. Users iteratively interact with the
% decision procedure, eliminating unreachable states from the invariant, moving
% towards the conditions under which \invariantclosure{} and
% \invariantconfluence{} are equivalent. Meanwhile, the decision procedure
% generates examples of reachable and unreachable states which help the user
% recognize patterns that describe reachability. We also leverage our intuitions
% about reachability to develop a new sufficient condition for
% \invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
% covers some cases that are not covered by \invariantclosure{}, and it can be
% automatically checked without any user interaction.

@@

Finally, we present Lucy: an implementation of our decision procedures and a
system for replicating \invariantconfluent{} and segmented
\invariantconfluent{} objects. Using Lucy, we find that our decision procedures
can efficiently handle a wide range of common workloads.  For example, in
\secref{Evaluation}, we apply Lucy to foreign key constraints and escrow
transactions. Lucy processes these workloads in less than half a second, and
no workload requires more than 66 lines of code to specify.
% Moreover, we find that segmented \invariantconfluent{} replication can achieve
% 10x to 100x more throughput than linearizable replication for low-coordination
% workloads.

% In closing, here is an outline of the paper and of our contributions:
% %
% We propose a novel expression-oriented definition of \invariantconfluence{}
% that is both formal and simple (\secref{InvariantConfluence}).
% %
% We develop an understanding of why \invariantclosure{} is not necessary for
% \invariantconfluence{} and use this understanding to develop conditions under
% which it is both necessary and sufficient (\secref{InvariantClosure}).
% %
% We exploit these conditions to design a general-purpose interactive decision
% procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
% %
% We again exploit these conditions to design a novel non-trivial sufficient
% condition for \invariantconfluence{}, called \mergereducibility.
% %
% We present segmented \invariantconfluence{}: a generalization of
% \invariantconfluence{} that uses a small amount of coordination to maintain
% invariants for replicated objects that are otherwise not \invariantconfluent{}
% (\secref{SegmentedInvariantConfluence}).
% %
% We evaluate our methods using a prototype implementation called Lucy
% (\secref{Evaluation}).

>>>>

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed
transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state
machine replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search} makes the application designer's life
very easy. To the developer, a strongly consistent system behaves exactly like
a single-threaded system running on a single node, so reasoning about the
behavior of the system is simple~\cite{herlihy1990linearizability}.
Unfortunately, strong consistency is at odds with performance. The CAP theorem
and PACELC theorem tell us that strongly consistent systems suffer from higher
latency at best and unavailability at worst~\cite{gilbert2002brewer,
brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
models like eventual consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak
consistency models strip an application developer of one of the earliest and
most effective tools that is used to reason about the execution of programs:
application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
if every transaction executing in a weakly consistent system individually
maintains an application invariant, the system as a whole can produce
invariant-violating states.

@@

Unfortunately, to date, the task of identifying whether or not an object
actually is \invariantconfluent{} has remained an exercise in human proof
generation. Bailis et al.\ manually categorized a set of common objects,
transactions, and invariants (e.g.\ foreign key constraints on relations,
linear constraints on integers) as \invariantconfluent{} or not. Hand-written
proofs of this sort are unreasonable to expect from programmers. Ideally we
would have a general-purpose program that could automatically determine
\invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
to make \invariantconfluence{} checkable:} to design a general-purpose
\invariantconfluence{} decision procedure, and implement it in an interactive
system.

@@

In this paper, we take a step back and study the underlying reason \emph{why}
\invariantclosure{} is not necessary for \invariantconfluence{}. Using this
understanding, we construct a set of modest conditions under which
\invariantclosure{} and \invariantconfluence{} are in fact \emph{equivalent},
allowing us to reduce the problem of determining \invariantconfluence{} to that
of determining \invariantclosure{}. Then, we use these conditions to design a
general-purpose interactive \invariantconfluence{} decision procedure.

@@

Finally, we present Lucy: an implementation of our decision procedures and a
system for replicating \invariantconfluent{} and segmented
\invariantconfluent{} objects. Using Lucy, we find that our decision procedures
can efficiently handle a wide range of common workloads. For example, in
\secref{Evaluation}, we apply Lucy to foreign key constraints and escrow
transactions. Lucy processes these workloads in less than half a second, and no
workload requires more than 66 lines of code to specify.
